\section{Experiments} \label{sect:experiments}

We have fully implemented and tested all methods% 
%
\footnote{All experiments on macOS 10.14.6, processor Intel Core i7 (2.6~GHz), RAM 16~GB 2400~MHz DDR4, OCaml 4.07.1. The OCaml runtime system options and garbage collection parameters are set as default.}$^,$%HACK!
%
\footnote{\textcolor{red}{The implementation is available: in GitHub?} \url{https://mega.nz/\#F!ec8y2ITa!tLOHKkHgs_7fjSI8m-O26Q}}
%
described in the previous sections. Our target language is untyped. Some dummies used for $pg$, $cpg$, $kpg$ and $xpg$ are replaced with the current updated views. This helps a program in the $put$ direction valid.

\subsection{Test Cases}

We have 
% conducted 
selected seven test cases (Table~\ref{tab:test-cases}) to 
represent
% assess 
non-trivial cases of practical significance.
The test cases use
% These $bx$s are 
left associative compositions because we focus on this kind of inefficiency in this paper.
% Although the number of cases is small, 
%% they are not trivial and can be completely used in practice. 
% they represent non-trivial cases of practical significance.
In the last two columns, $s$ and $v$ are the updated source and view, respectively. They are produced by applying $put$ and $get$ to the original source $s_0$ and view $v_0$. That is, $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$, where $bx$ is the program indicated in the second column of the table. 
% are only interested in this kind of programs in the paper. 
% Please note that the results
Results $s$ and $v$ are independent of the associativity 
% style 
of the composition.

\begin{table}[hbt!]
    \centering
    \caption{Composition test cases (number of compositions = $n$)}
    \label{tab:test-cases}
    \begin{tabular*}{\textwidth}{|l @{\extracolsep{\fill}}|l|l|c|c|c|c|}
        \hline
        \multirow{2}{*}{No} & \multicolumn{1}{c|}{\multirow{2}{*}{Name}} & \multicolumn{1}{c|}{\multirow{2}{*}{Type}} & \multicolumn{2}{c|}{Input} & \multicolumn{2}{c|}{Output} \\ \cline{4-7} 
        & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{$s_0$} & \multicolumn{1}{c|}{$v_0$} & \multicolumn{1}{c|}{$s$} & \multicolumn{1}{c|}{$v$} \\ \hline
        1 & lcomp-phead-ldata & straight line & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
        2 & lcomp-ptail & straight line & [1,$\ldots$,n+1] & [1,$\ldots$,10] & $s_0 \ @ \ v_0$ & [\ ] \\ \hline
        3 & lcomp-ptail-ldata & straight line & $\underbrace{[L,\ldots,L]}_{\text{n+1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{10 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n+11 times}}$ & [\ ] \\ \hline
        4 & lcomp-bsnoc & straight line & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] \\ \hline
        5 & lcomp-bsnoc-ldata & straight line & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]]}_{\text{n-1 times}}$ \\ \hline
        6 & breverse & recursion & [1,$\ldots$,n] & [1,$\ldots$,n] & [n,$\ldots$,1] & [n,$\ldots$,1] \\ \hline
        7 & breverse-ldata & recursion & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ \\ \hline
    \end{tabular*}
\end{table}

% The first 5 cases merely use $n$ $\circ$ operators to make straight-line compositions from the same $n + 1$ \textit{pure} programs. 
The first five test cases
(1--5) 
are $n$ straight-line (non-recursive) compositions of the same $n + 1$ 
% non-recursive 
programs.
% We refer to them as `pure' compositions.
% We use the term \textit{pure} here to refer non-recursion compositions. 
The prefix \textit{lcomp} in the name of a test case indicates that the textual compositions are left associative.
% that the associate precedence of $\circ$ in a straight-line composition is left.
The suffix \textit{ldata} indicates that the input size is considered large. The symbol $L$ in the input column 
% \textit{ldata} test cases 
% can be expressed as a complex list as follows: 
stands for a list $L = [T,\ldots,T]$ with
% $(n \ T\text{s}), 
$T = [A,\ldots,A]$ of length 10
% \ (10 \ A\text{s}),
and
$A = [1,\ldots,5]$. 
% These structures are 
They are only intended to generate test data that is large enough for 
% observing 
measuring
% experimental 
results. 

We introduced the composition $phead \circ phead$ earlier in Section~\ref{sec:intro}. The composition of many \textit{phead}s 
% do things
works similarly. The head of a head element inside a 
%super 
deeply nested list, which 
% expresses 
is the source, 
% should be 
is updated by the changed view. Because of the type of the source, this program is 
% seen 
categorized as a \textit{ldata} case. 

Next, we briefly explain the behavior of the remaining compositions in 
\mbox{Table\,\ref{tab:test-cases}.}%HACK!

The composition of many $ptail$s, in the \textit{put} direction, replaces a part of the tail of the source list by the view list and, in the \textit{get} direction, returns such a tail from the source.

The composition of many \textit{bsnoc}s, in the \textit{put} direction, creates a permutation of the view list if its length is not larger than the length of the source list and, in the \textit{get} direction, produces a permutation of the source list.

\textit{breverse} is defined in terms of \textit{bfoldr}, that appeared in Section \ref{sect:minbigul}. In the \textit{put} direction, it produces a reverse of the view list if its length is not larger than the length of the source list and, in the \textit{get} direction, produces a reverse of the source list. Note that compositions are by the recursions of \textit{breverse} and 
the number of compositions are dynamically determined by the length of the source list.

\subsection{Results}

\input{result.tex}


Figure \ref{fig:result} 
% illustrates 
shows the evaluation times for each of the seven test cases using the three methods: $put$ in minBiGUL, $put_m$ in minBiGUL$_m$ and $xpg$.
We also did similar experiments with $pg$, $cpg$ and $kpg$, but their results are slower than the corresponding ones of $xpg$. The slowness was caused by the exponential number of $pg$ calls in the case of $pg$, redundant evaluations ($cpg$), and redundant overhead for constructing closures ($kpg$).
Therefore we simply omit these results in Figure \ref{fig:result}.
As we all know, $put$ works poorly for left associative compositions because of the number of reevaluated $get$s.
%The figure shows $put_m$ and $xpg$ are better than $put$ in some cases.
The left part of the figure contains tests using not-large inputs, and we see that $put_m$ is the fastest method for them. However if the input size is large enough, in the cases of the right part of the figure, $put_m$ will be slower quickly due to time for manipulating data in the table. At that time, $xpg$ is 
% proved to be
the most effective method.

\textcolor{red}{Note that this result focuses on the BX programs that includes many compositions. If the number of compositions are small in the BX programs, $put$ will be fastest, because of the overhead for memoization ($put_m$) and the overhead for keeping complements in closures ($xpg$).}

\textcolor{red}{TODO-B2: add short evaluation result of memory.}
